---
title: "p8105_hw6_tt3022"
author: "Thomas Tang"
date: "2024-12-02"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(modelr)   
```

Problem 2
```{r warning=FALSE}
homicide_data <- read_csv("data/homicide-data.csv", locale = locale(encoding = "latin1"))

homicide_data <- homicide_data %>%
  mutate(city_state = paste(city, state, sep = ", "))


homicide_data <- homicide_data %>%
  mutate(solved = ifelse(grepl("Closed", disposition), 1, 0))


omit_cities <- c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL")
filtered_data <- homicide_data %>%
  filter(!(city_state %in% omit_cities))


filtered_data <- filtered_data %>%
  filter(victim_race %in% c("White", "Black"))


filtered_data <- filtered_data %>%
  mutate(victim_age = as.numeric(victim_age)) %>%
  filter(!is.na(victim_age))
```
I deleted omit cities and remove the unknown value in victim_age


Then I fit a logistic regression model
Outcome: resolved (1 = solved, 0 = unsolved)
Predictors: victim_age, victim_sex, and victim_race
```{r}
baltimore_data <- filtered_data %>%
  filter(city_state == "Baltimore, MD")

fit_logistic <- glm(solved ~ victim_age + victim_race + victim_sex, 
                    data = baltimore_data, 
                    family = binomial())


logistic_results <- summary(fit_logistic)$coefficients %>%
  as.data.frame() %>%
  mutate(
    term = rownames(.),
    OR = exp(Estimate),  # Odds Ratios
    lower_CI = exp(Estimate - 1.96 * `Std. Error`),  # Lower 95% CI
    upper_CI = exp(Estimate + 1.96 * `Std. Error`)   # Upper 95% CI
  ) %>%
  select(term, Estimate, OR, lower_CI, upper_CI, `Pr(>|z|)`)


write.csv(logistic_results, "logistic_results_baltimore.csv", row.names = FALSE)
logistic_results
```
The odds ratio for victim_sexMale is 0.35, which means the odds of solving a homicide for male victims are 65% lower compared to female victims, after adjusting for age and race.


```{r}
results <- filtered_data %>%
  group_by(city_state) %>%
  nest() %>%
  mutate(
    model = map(data, ~ glm(solved ~ victim_sex + victim_age + victim_race, 
                            data = ., 
                            family = binomial())),
    or_ci = map(model, ~ {
      
      coef_summary <- summary(.)$coefficients
      log_odds <- coef_summary["victim_sexMale", "Estimate"]
      std_error <- coef_summary["victim_sexMale", "Std. Error"]
      or <- exp(log_odds)
      lower_ci <- exp(log_odds - 1.96 * std_error)
      upper_ci <- exp(log_odds + 1.96 * std_error)
      tibble(
        OR = or,
        lower_CI = lower_ci,
        upper_CI = upper_ci,
        p_value = coef_summary["victim_sexMale", "Pr(>|z|)"]
      )
    })
  ) %>%
  select(city_state, or_ci) %>%
  unnest(cols = or_ci)



```

create plot
```{r}
results <- results %>%
  arrange(OR)  


ggplot(results, aes(x = OR, y = reorder(city_state, OR))) +
  geom_point(size = 3, color = "orange") +
  geom_errorbarh(aes(xmin = lower_CI, xmax = upper_CI), height = 0.2, color = "orange") +
  geom_vline(xintercept = 1, linetype = "dashed", color = "red") +
  labs(
    title = "Adjusted Odds Ratios for Male vs Female Homicide Solving by City",
    x = "Odds Ratio (Log Scale)",
    y = "City"
  ) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 10),
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    plot.title = element_text(hjust = 0.5, size = 14)
  )
```


Problem 3:
```{r}

```

```{r}
birthweight_data <- read.csv("data/birthweight.csv")

# View the structure of the dataset
str(birthweight_data)

# Convert appropriate numeric variables to factors
birthweight_data <- birthweight_data %>%
  mutate(
    babysex = factor(babysex, labels = c("Male", "Female")),
    frace = factor(frace, levels = c(1, 2, 3, 4, 8), 
                   labels = c("White", "Black", "Asian", "Puerto Rican", "Other")),
    mrace = factor(mrace, levels = c(1, 2, 3, 4, 8), 
                   labels = c("White", "Black", "Asian", "Puerto Rican", "Other")),
    malform = factor(malform, levels = c(0, 1), 
                     labels = c("Absent", "Present"))
  )

# Check for missing values
missing_data_summary <- colSums(is.na(birthweight_data))
print(missing_data_summary)
```



```{r}
model <- lm(bwt ~ gaweeks + delwt + momage + babysex + smoken, data = birthweight_data)


birthweight_data <- birthweight_data %>%
  add_predictions(model, var = "fitted") %>%
  add_residuals(model, var = "residuals")


ggplot(birthweight_data, aes(x = fitted, y = residuals)) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(
    title = "Residuals vs Fitted Values",
    x = "Fitted Values",
    y = "Residuals"
  ) +
  theme_minimal()
```
A linear regression model was used because the outcome variable (bwt) is continuous, and the relationship between the predictors and outcome was assumed to be linear.
The regression model was fitted using the formula:
bwt ~ gaweeks + delwt + momage + babysex + smoken
Randomness of residuals around zero (indicates good fit).

comparing:

```{r}
# Define the formulas for the three models
formulas <- list(
  model1 = bwt ~ gaweeks + delwt + momage + babysex + smoken,
  model2 = bwt ~ blength + gaweeks,
  model3 = bwt ~ bhead * blength * babysex  # Includes interactions
)

# Perform Monte Carlo cross-validation
set.seed(123)  # For reproducibility
cv_results <- crossv_mc(birthweight_data, n = 10) %>%
  mutate(
    model1_mse = map2_dbl(train, test, ~ {
      mod <- lm(formulas$model1, data = as.data.frame(.x))
      preds <- predict(mod, newdata = as.data.frame(.y))
      mean((as.data.frame(.y)$bwt - preds)^2)
    }),
    model2_mse = map2_dbl(train, test, ~ {
      mod <- lm(formulas$model2, data = as.data.frame(.x))
      preds <- predict(mod, newdata = as.data.frame(.y))
      mean((as.data.frame(.y)$bwt - preds)^2)
    }),
    model3_mse = map2_dbl(train, test, ~ {
      mod <- lm(formulas$model3, data = as.data.frame(.x))
      preds <- predict(mod, newdata = as.data.frame(.y))
      mean((as.data.frame(.y)$bwt - preds)^2)
    })
  )

# Summarize the results
cv_summary <- cv_results %>%
  summarise(
    model1_mean_mse = mean(model1_mse),
    model2_mean_mse = mean(model2_mse),
    model3_mean_mse = mean(model3_mse)
  )
print(cv_summary)

```
Model 3 performs the best in terms of predictive accuracy with the lowest MSE.
Model 2 performs better than Model 1, showing that length and gestational age are significant predictors, even in a simpler setup.




